---
title: "Description of the data"
description: |
  In this post, we will discuss how we select predictors used in our model to predicor individual math score. In addition to the original training set, we selected and merged a few extra predictors that are at higher levels, including staff info, school characteristic, lunch eligibility, directory and membership. The exact procedure and rationale behind choosing those predictors are discussed below. \
  Note that this post only discuss the procedure by which we build up our training set. This post does not include details on feature enginerring. The details of building recipes can be found in each prelim fit post individually. 
author:
  - name: Peeta Li, Yufei Zhao, Bernice Cheung
    url: https://example.com/norajones
date: 06-07-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rio)
library(vip)
library(skimr)
library(janitor)
```

## Create a Merged Dataset

### The original training data set 
First, we read in the original training set. We then \
1. Made sure that the 'ncessch' variable is numeric, which will be used for joinning two data sets. \
2.Removed the classification variable because we are going to build a regression model. \
3.Changed the NA in variables 'lang_cd' and 'ayp_lep' to something meaningful. We did this here instead of in the recipe becuase the NAs do have meaning, instead of being unknown. For example, the NAs in 'lang_cd' suggests that one's first language is English, otherwise Spanish. 

The meanings of each variable in the original training dataset listed below.

```{r}
# Original training set 
train <- import(here::here('data','train.csv')) %>% 
  mutate(ncessch = as.double(ncessch)) %>% 
  select(-classification) %>%
  mutate(lang_cd = ifelse(is.na(lang_cd), "E", lang_cd),
         ayp_lep = ifelse(is.na(ayp_lep), "G", ayp_lep))

dic <- import(here::here('data', 'data_dictionary.csv'))
knitr::kable(dic)
```

Next we imported multiple extra datasets from <https://nces.ed.gov/ccd/files.asp#Fiscal:2,LevelId:7,SchoolYearId:32,Page:1>. We extracted certain predictors from these dataset and merged them to the original training set. The details and rationales are discribed below. 

### The Staff Dataset
Through reading the dictionaries on the varaible names (i.e., documentations), we choose the number of teachers in the STAFF data set. We also select 'ncessch' for data joining purpose. 

```{r}
# add in extra data 
# 1) staff
staff <- import(here::here('data','extra_data','staff','ccd_sch_059_1718_l_1a_083118.csv')) %>%
  filter(ST == "OR")  %>%
  select(NCESSCH, TEACHERS) %>% # Teachers: number of teachers (numeric)
  janitor::clean_names() %>%
  mutate(ncessch = as.double(ncessch))

```

We then left join the original training set (i.e., train) with the selected staff data set by the variable 'ncessch'. We repeat this left join step for each data set. Thus the repeated step would not be specified below. 

```{r}
# join train and staff by ncessch
df <- train %>% 
  left_join(staff)
```

### The School Characteristic Dataset
Merely reading the documentation on this dataset did not help us make sense of which variable we should select. Thus, we made some plots to visualize which categorical predictors in the school-characteristic dataset have influence on individual math score. That is, the averaged math scores differ across different groups in a given categorical predictor.

```{r}
# 2) scool characteristics

# clean the school characteristic dataset
schoolDf <- import(here::here('data','extra_data','school_characteristics','ccd_sch_129_1718_w_1a_083118.csv'))%>% 
  filter(ST == "OR") %>% 
  janitor::clean_names() %>% 
  mutate(ncessch = as.double(ncessch)) %>%
  modify_if(is.character, as.factor)

# Merge with the full dataset
fullDf <- left_join(train, schoolDf, by= "ncessch") %>%
  select(score, colnames(schoolDf))

# Plot bar graph to explore the data
expl <- names(fullDf[,2:21])
# func for map
explore_fun <- function(x, y) {
    ggplot(fullDf, aes(x = .data[[x]], y = .data[[y]]) ) +
    geom_bar(stat="identity") + 
    theme_minimal()
}
elev_plots <- map(expl, ~explore_fun(.x, "score"))
```

We explored all variables in the School Characteristic Dataset, stored in the 'elev_plots' variable the chunk above. For demonstration purpose, we only showed one barplot for one of the variables we chose. 

We chose to select the variable 'nslp_status' because the mean of individual math scores differ across different groups in this variable, as illustrated below.

```{r}
elev_plots[17]
```

Given the criterion discribed above, we eventually chose variables 'TITLE1_STATUS', indicating title 1 status, 'NSLP_STATUS', indicating national school lunch program and 'Virtual', indicating school virtual status

```{r}
school_char <- import(here::here('data','extra_data','school_characteristics','ccd_sch_129_1718_w_1a_083118.csv')) %>%
  filter(ST == "OR")  %>%
  select(NCESSCH, TITLEI_STATUS, NSLP_STATUS, VIRTUAL) %>% 
  janitor::clean_names()  %>% 
  mutate(ncessch = as.double(ncessch))

# join df and school_char by ncessch
df <- df %>% 
  left_join(school_char)

```

### The Lunch Eligibility data set
Here, we basically followed the feature selection procedure we were shown in class, selecting the type of lunch program (i.e., LUNCH_PROGRAM) and the number of student in each lunch program (i.e.,STUDENT_COUNT). We then transformed the data form from long to wide, thus each unique ncessch id has its own row with different columns being differnt lunch programs and the cells are the counts of students. Followed the class procedure, from online dataset we extracted the number of student in each ncessch. Using the two student count, we calculated the proportion of free lunch (i.e., free_lunch_prp) and the proportion of reduced lunch (i.e., reduced_lunch_prop)

```{r}
# 3) lunch eligibility
lunch_elig <- import(here::here('data','extra_data','lunch_program_eligibility','ccd_sch_033_1718_l_1a_083118.csv')) %>%
  filter(ST == "OR")  %>%
  select(NCESSCH, LUNCH_PROGRAM, STUDENT_COUNT)  %>%
  janitor::clean_names()  %>% 
  select(ncessch, lunch_program, student_count)  %>% 
  mutate(student_count = replace_na(student_count, 0))  %>% 
  pivot_wider(names_from = lunch_program,
              values_from = student_count)  %>% 
  janitor::clean_names()  %>% 
  mutate(ncessch = as.double(ncessch))

stu_counts <- import("https://github.com/datalorax/ach-gap-variability/raw/master/data/achievement-gaps-geocoded.csv",
                     setclass = "tbl_df")  %>% 
  filter(state == "OR" & year == 1718)  %>% 
  count(ncessch, wt = n)  %>% 
  mutate(ncessch = as.double(ncessch))

lunch_elig <- left_join(lunch_elig, stu_counts)

lunch_elig <- lunch_elig %>% 
  mutate(free_lunch_prop = free_lunch_qualified / n,
         reduced_lunch_prop = reduced_price_lunch_qualified / n)  %>% 
  select(ncessch, ends_with("prop"))

df <- df %>% 
  left_join(lunch_elig)
  
```

### Directory dataset 
Through reading the documentation for the dataset, we choose the varaibles indicating the type of school (i.e., SCH_TYPE_TEXT) and whether the school is a charter school (i.e., CHARTER_TEXT)

```{r}
# 4) directory
directory <- import(here::here('data','extra_data','directory','ccd_sch_029_1718_w_1a_083118.csv')) %>%
  filter(ST == "OR")  %>%
  select(NCESSCH, SCH_TYPE_TEXT, CHARTER_TEXT) %>% 
  janitor::clean_names() %>% 
  mutate(ncessch = as.double(ncessch))

df <- df %>% 
  left_join(directory)
```
### Membership dataset
Here, we basically followed the feature selection procedure we were shown in class, selecting the school name (i.e., sch_name) and the two Ids for joinning data. 

```{r}
# 5) membership
sheets <- readxl::excel_sheets(here::here('data','extra_data','fallmembershipreport_20192020.xlsx'))
ode_schools <- readxl::read_xlsx(here::here('data','extra_data','fallmembershipreport_20192020.xlsx'),
                                 sheet = sheets[4])
ethnicities <- ode_schools %>% 
  select(attnd_schl_inst_id = `Attending School ID`,
         attnd_dist_inst_id = `Attending District Institution ID`,
         sch_name = `School Name`,
         contains('%')) %>%
  janitor::clean_names()
names(ethnicities) <- gsub('x2019_20_percent', 'p', names(ethnicities))
```

Here we join the membership dataset with the merged df dataset by the two Ids. 

```{r}
df <- df %>% 
  left_join(ethnicities) 
```

## Explore the Merged Dataset

We first changed all character variables into factors and then used {skimr} to get a sense of all predictors we have in the data set.

```{r}
df <- df%>%
  modify_if(is.character, as.factor)
(df_skim <- skim(df))
```

After skimming the data, we noticed that the variable 'calc_admn_cd' has all NAs, thus we removed this varaible. The remaining dataframe includes in total 56 predicors, among which 34 are categorical and 22 are numerical. 

```{r}
df <-  df %>%
  select(-calc_admn_cd)
```

Then we decided to have a closer look to the distribution of all the categorical variables using {janitor}

```{r}
factor_vars <- 
  df_skim %>%
  filter(skim_type == 'factor') %>%
  select(skim_variable) 

factor_vars$skim_variable[-34] %>%
  map(~tabyl(df,.))
```

Last, we decided to use {vip} to explore the importance for each predictor. To do this we need to first remove all NAs, and categorical predictors that only have one level. 

```{r}
df_forvip <- df %>%
  drop_na()

df_forvip <-  df_forvip  %>% 
  select(-ayp_dist_partic , -rc_dist_partic, -grp_rpt_dist_partic)

#vip(lm(score ~ ., df_forvip), mapping = aes(fill = Sign))
```

We saved this merged dataframe as our training set. 

```{r, eval = F}
write.csv(df, 'final_merged_train.csv', row.names = F)
```

The exact same procedure is be applied to the testing set, for which we will generate the predictions using our finalized model. 

```{r, eval = F}
test <- import(here::here('data','test.csv')) %>% 
  mutate(ncessch = as.double(ncessch)) %>% 
  mutate(lang_cd = ifelse(is.na(lang_cd), "E", lang_cd),
         ayp_lep = ifelse(is.na(ayp_lep), "G", ayp_lep))

# join staff dataser
test_full <- test %>% 
  left_join(staff)

# join school characteristic dataset
test_full <- test_full %>% 
  left_join(school_char)

# join lunch eligibility dataset
test_full <- test_full %>% 
  left_join(lunch_elig)

# join directory dataset
test_full <- test_full %>% 
  left_join(directory)

# join membership dataset
test_full <- test_full %>% 
  left_join(ethnicities) 

test_full <- test_full%>%
  modify_if(is.character, as.factor)
test_full <- test_full %>%
  select(-calc_admn_cd)

write.csv(test_full, here::here('data','final_merged_test.csv'), row.names = F)

```
